# Exploration

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
taxi_data <- read.csv('~/Desktop/2021_taxi_data.csv')
```
My goal is to look into the factors that determine how people tip. To do this, I want to create a "tipped" column that is just a binary of whether or not somebody tipped at all. I also want to create a "tip" column that shows whether or not somebody tipped over 10% of the fare amount. 

Note: initially, I tried to use 15% as the threshold. After doing some analysis on results and thinking about them, I decided that 10% would be a better threshold as it would indicate one tipping above a negligable amount. Usually, 15% is a pretty standard tip for some who meant to tip at all. 

```{r}
library(ggplot2)
colSums(is.na(taxi_data))
```
To understand the columns and what the values mean, I used this site:
https://www.nyc.gov/assets/tlc/downloads/pdf/data_dictionary_trip_records_yellow.pdf

It seems that most columns don't have NAs. The ones that do are passenger_count, VendorID, RatecodeID, payment_type, and congestion_surcharge. 
Passenger_count: just remove the rows that don't have this value
VendorID: replace NA with 0 to represent unknown
RatecodeID: not sure
payment_type: replace all NA with code 5 (unknown)
congestion_surcharge: not sure

```{r}
taxi_data["VendorID"][is.na(taxi_data["VendorID"])] <- 0
taxi_data["payment_type"][is.na(taxi_data["payment_type"])] <- 5
# taxi_data <- taxi_data[!is.na(taxi_data$passenger_count),]
colSums(is.na(taxi_data))
```
For some reason, removing all rows with passenger_count NA removes all other NA instances. This is good!

```{r}
# create tip column
tipped <- ifelse(taxi_data$tip_amount == 0, 0, 1)
taxi_data$tipped <- tipped
mean(!taxi_data$tipped)

taxi_data <- taxi_data[taxi_data$fare_amount != 0, ]
dim(taxi_data)

tip_percent <- taxi_data$tip_amount / taxi_data$fare_amount
taxi_data$tip_percent <- tip_percent

tip <- ifelse(taxi_data$tip_percent < 0.15, 0, 1)
taxi_data$tip <- tip
mean(!taxi_data$tip)
# of the sampled data, 29.6% of rides were tipped under 10%.
```

As we can see, of the sampled data, 27.1% of rides were NOT tipped and 29.6% of rides were tipped under 10% (including not tipped at all). Next, I want to get a general sense of the distribution of tip amounts and the distribution of fare amounts. 

```{r}
hist(taxi_data$fare_amount, main="Fare Amount", xlab="fare", breaks=20)
negative <- taxi_data[taxi_data$fare_amount < 0, ]
```
This is odd- there seem to be some rides where the fare amount is negative (the taxi driver owes the passenger the fare amount). I did some research and attributed this to messy data, so I removed the negative fares and redid the histogram. 

```{r}
taxi_data <- taxi_data[taxi_data$fare_amount >= 0, ]
hist(taxi_data$fare_amount, main="Fare Amount", xlab="fare", breaks=40)
# for random forest, you need strings as factors (read it in as "string as factor" when reading the CSV)
# this is a problem because random forest will do REGRESSION instead of CLASSIFICATION
# stringsAsFactors = TRUE
```
We see that most fares seem to be around \$5-20. I'd like to make a categorical variable for rides over and under \$20.

```{r}
taxi_data$high_fare <- ifelse(taxi_data$fare_amount < 20, 0, 1)
mean(taxi_data$high_fare)
```

So around 14.4% of all rides are over $20.

```{r}
hist(taxi_data$tip_amount, main="Tip Amount", xlab="tip", breaks=20)
```
We see a spike at 0 representing the proportion of rides that weren't tipped at all. Generally, we see that most rides are tipped between the $0-5 range. 

```{r}
hist(taxi_data$tip_percent, main="Tip Percentage", xlab="tip", xlim = c(0, 1), breaks=100)
```

We see a spike at 0 representing the proportion of rides that aren't tipped at all. We also see a spike around 25-30% tip. 

Next, I want to see how long these rides last. To do that, I need to first convert the pickup/dropoff times to time variables, and find the difference. 

```{r}
taxi_data$pickup <- as.POSIXct(taxi_data$tpep_pickup_datetime, format = "%m/%d/%Y %I:%M:%S %p")
taxi_data$dropoff <- as.POSIXct(taxi_data$tpep_dropoff_datetime, format = "%m/%d/%Y %I:%M:%S %p")

# minutes was the most reasonable unit of measurement for ride duration
duration <- difftime(taxi_data$dropoff, taxi_data$pickup, units = 'mins')
taxi_data$duration <- duration

taxi_data <- taxi_data[taxi_data$duration < 120, ]
```

In my exploration of the data, there were some values that seemed like outliers (rides that took multiple hours). I removed 11 of these rows from the data set. 

Next, I want to create a categorical variable for the season in which the ride took place based on the pickup time. To do this, I will looked into this package: hydroTSM. Maybe the season in which the ride took place will impact the amount riders tip!


```{r}
library("hydroTSM")
season <- time2season(taxi_data$pickup, out.fmt = "seasons")
taxi_data$season <- season
```

Finally, I would like to create a categorical variable that takes into account the pickup and dropoff locations. I don't want it to be too granular, so I'm using the "pulocationID" and "dolocationID" columns. Just looking at the numbers, they don't mean anything so I looked into what the IDs mean. I found this dataset: https://catalog.data.gov/dataset/nyc-taxi-zones and named it pickup_ids in my code. I cross-referenced the "borough" column of pickup_ids to figure out the borough each pickup ID corresponds to. 

This link was also helpful. https://www.nyc.gov/assets/tlc/downloads/pdf/data_dictionary_trip_records_yellow.pdf

```{r}
zone_ids <- read.csv('data/taxi_zones.csv')
# zone_ids

pickup_borough <- c()
dropoff_borough <- c()

for (i in 1:nrow(taxi_data)) {
  pickup_id <- taxi_data$PULocationID[i]
  dropoff_id <- taxi_data$DOLocationID[i]
  
  pu_id <- zone_ids[zone_ids$LocationID == pickup_id, ]
  pu_id
  pu_borough <- pu_id$borough
  # print(pu_borough)
  if (length(pu_borough) == 0) {
    pickup_borough[i] <- "unknown"
  }
  else {
    pickup_borough[i] <- pu_borough
  }
  
  
  do_id <- zone_ids[zone_ids$LocationID == dropoff_id, ]
  do_borough <- do_id$borough
  if (length(do_borough) == 0) {
    dropoff_borough[i] <- "unknown"
  }
  else {
    dropoff_borough[i] <- do_borough
  }
}

table(pickup_borough)
table(dropoff_borough)

taxi_data$pickup_borough <- pickup_borough
taxi_data$dropoff_borough <- dropoff_borough
```

Let's look at the relative frequencies of the passenger count.

```{r}
hist(taxi_data$passenger_count, main="Passenger Count", xlab="passenger count", breaks=20)
```

For the most part, we have rides with only 1 passenger, followed by 2. What's surprising to me is that there are rides with 5-6 passengers at all; over covid, I remember that most taxis only allowed 1-3 passengers to ride but maybe by late 2021 they lifted the restrictions. 

```{r}
taxi_data$passenger_count <- factor(taxi_data$passenger_count)
table(taxi_data$passenger_count)
ggplot(taxi_data, aes(x=passenger_count, y=tip_percent)) +
  geom_boxplot(outlier.shape = NA) +
  ylim(0, 1) +
  coord_flip()
```
This is interesting- there doesn't seem to be much of a trend in passenger count vs tip percentage. 

```{r}
taxi_data$pickup_borough <- factor(taxi_data$pickup_borough)
table(pickup_borough)
ggplot(taxi_data, aes(x=pickup_borough, y=tip_percent)) +
  geom_boxplot(outlier.shape = NA) +
  ylim(0, 1) +
  coord_flip()
```

```{r}
library(forcats)
ggplot(taxi_data, aes(fct_infreq(dropoff_borough), fill=factor(tip))) +
  geom_bar(position="dodge")
```
At this point in my iterations, I've created all the new categorical variables I'd like to create for my data analysis. 

```{r}
taxi_data$tip <- factor(taxi_data$tip)
write.csv(taxi_data, file='data/2021_taxi_data_mod.csv')
```




