# [Exploration]

```{r}
taxi_data <- read.csv('~/Desktop/2021_taxi_data.csv')
head(taxi_data)
```
My goal is to look into the factors that determine how people tip. To do this, I want to create a "tipped" column that is just a binary of whether or not somebody tipped at all. I also want to create a "tip" column that shows whether or not somebody tipped over 15% of the fare amount. 

```{r}
# create tip column
tipped <- ifelse(taxi_data$tip_amount == 0, 0, 1)
taxi_data$tipped <- tipped
mean(!taxi_data$tipped)

taxi_data <- taxi_data[taxi_data$fare_amount != 0, ]
dim(taxi_data)

tip <- ifelse(taxi_data$tip_amount / taxi_data$fare_amount < 0.15, 0, 1)
taxi_data$tip <- tip
mean(!taxi_data$tip)
# of the sampled data, 35.2% of rides were tipped under 15%.
```

As we can see, of the sampled data, 27% of rides were NOT tipped and 35.2% of rides were tipped under 15% (including not tipped at all). Next, I want to get a general sense of the distribution of tip amounts and the distribution of fare amounts. 

```{r}
hist(taxi_data$tip_amount, main="Tip Amount", xlab="tip", breaks=20)
```
We see a spike at 0 representing the proportion of rides that weren't tipped at all. Generally, we see that most rides are tipped between the $0-5 range. 

```{r}
hist(taxi_data$tip_amount / taxi_data$fare_amount, main="Tip Percentage", xlab="tip", xlim = c(0, 1), breaks=100)
```

We see a spike at 0 representing the proportion of rides that aren't tipped at all. We also see a spike around 25-30% tip. 

```{r}
hist(taxi_data$fare_amount, main="Fare Amount", xlab="fare", breaks=20)
negative <- taxi_data[taxi_data$fare_amount < 0, ]
negative
```
This is odd- there seem to be some rides where the fare amount is negative (the taxi driver owes the passenger the fare amount). I did some research and attributed this to messy data, so I removed the negative fares and redid the histogram. 

```{r}
taxi_data <- taxi_data[taxi_data$fare_amount >= 0, ]
hist(taxi_data$fare_amount, main="Fare Amount", xlab="fare", breaks=40)
# for random forest, you need strings as factors (read it in as "string as factor" when reading the CSV)
# this is a problem because random forest will do REGRESSION instead of CLASSIFICATION
# stringsAsFactors = TRUE
```
We see that most fares seem to be around $5-20.

Next, I want to see how long these rides last. To do that, I need to first convert the pickup/dropoff times to time variables, and find the difference. 

```{r}
taxi_data$pickup <- as.POSIXct(taxi_data$tpep_pickup_datetime, format = "%m/%d/%Y %I:%M:%S %p")
taxi_data$dropoff <- as.POSIXct(taxi_data$tpep_dropoff_datetime, format = "%m/%d/%Y %I:%M:%S %p")

# minutes was the most reasonable unit of measurement for ride duration
duration <- difftime(taxi_data$dropoff, taxi_data$pickup, units = 'mins')
taxi_data$duration <- duration

taxi_data <- taxi_data[taxi_data$duration < 120, ]
```

In my exploration of the data, there were some values that seemed like outliars (rides that took multiple hours). I removed 11 of these rows from the data set. 

Next, I want to create a categorical variable for the season in which the ride took place based on the pickup time. To do this, I will looked into this package: hydroTSM. Maybe the season in which the ride took place will impact the amount riders tip!

```{r}
library("hydroTSM")
season <- time2season(taxi_data$pickup, out.fmt = "seasons")
taxi_data$season <- season

```

Finally, I would like to create a categorical variable that takes into account the pickup and dropoff locations. I don't want it to be too granular, so I'm using the "pulocationID" and "dolocationID" columns. Just looking at the numbers, they don't mean anything so I looked into what the IDs mean. I found this dataset: https://catalog.data.gov/dataset/nyc-taxi-zones and named it pickup_ids in my code. I cross-referenced the "borough" column of pickup_ids to figure out the borough each pickup ID corresponds to. 

This link was also helpful. https://www.nyc.gov/assets/tlc/downloads/pdf/data_dictionary_trip_records_yellow.pdf

```{r}
zone_ids <- read.csv('~/Desktop/taxi_zones.csv')
# zone_ids

pickup_borough <- c()
dropoff_borough <- c()

for (i in 1:nrow(taxi_data)) {
  pickup_id <- taxi_data$PULocationID[i]
  dropoff_id <- taxi_data$DOLocationID[i]
  
  pu_id <- zone_ids[zone_ids$LocationID == pickup_id, ]
  pu_id
  pu_borough <- pu_id$borough
  # print(pu_borough)
  if (length(pu_borough) == 0) {
    pickup_borough[i] <- "unknown"
  }
  else {
    pickup_borough[i] <- pu_borough
  }
  
  
  do_id <- zone_ids[zone_ids$LocationID == dropoff_id, ]
  do_borough <- do_id$borough
  if (length(do_borough) == 0) {
    dropoff_borough[i] <- "unknown"
  }
  else {
    dropoff_borough[i] <- do_borough
  }
}

table(pickup_borough)
table(dropoff_borough)

taxi_data$pickup_borough <- pickup_borough
taxi_data$dropoff_borough <- dropoff_borough
```

```{r}
taxi_data$passenger_count <- factor(taxi_data$passenger_count)
table(passenger_count)
ggplot(taxi_data, aes(x=passenger_count, y=tip_amount)) +
  geom_boxplot() +
  coord_flip()
```

At this point in my iterations, I've created all the new categorical variables I'd like to create for my data analysis. 

```{r}
write.csv(taxi_data, file='~/Desktop/2021_taxi_data_mod.csv')
```




